Wigglyhedra

### 2024-06-12 draft/wigglyhedra.tex
Minor readability updates.

We now describe the \(t\)-structure induced by the full subcategory of \emph{linear complexes} of \(\mathcal{C}_m\); a variant of this construction is described in~\cite[Section 2.5]{lic.que:21}.
It is convenient to think of an object \(P^{\bullet}\) of \(\mathcal{C}_m\)
as bigraded (projective) \(A_m\)-module equipped with a differential \(d \colon P^{\bullet} \to P^{\bullet}\).
More precisely, we can write
\[P^{\bullet} = \bigoplus_{g,h,i} M_i\langle g \rangle[h],\]
where \(g\) is the internal grading shift, and \(h\) is the homological grading shift, and 
each object \(M_i\) is a direct sum of copies of objects in \(\{P_1,\ldots, P_m\}\).
Set \(P^h\) to be the direct sum of all summands that have homological grading shift \(h\).
That is, in the notation above, we have
\[P^{h} = \bigoplus_{g,i} M_i\langle g \rangle[h].\]

@@

The Khovanov--Seidel correspondence (explained in Section 4 of~\cite{kho.sei:02}) associates objects of \(\mathcal{C}_{n+1}\) (up to grading\asilata{Should this say homological?} shift) to curves on a configuration of \(n+2\) marked points arranged in a straight line.

@@

\begin{lemma}
  Every wiggly arc corresponds to an object in the heart \(\mathcal{A}_m\) of the standard \(t\)-structure on \(\mathcal{C}_m\).
\end{lemma}
\begin{proof}
  Let \(P^{\bullet}\) be the complex built from an arbitrary (not necessarily wiggly) arc.
  The procedure to build up the complex of projective modules from a curve explicitly describes the differentials.

  Each component of these differentials is a map of the form
  \[P_i\langle g \rangle[h] \to P_j\langle g' \rangle[h-1].\]
  First consider a component of the differential from some \(P_i\langle g \rangle\) to some \(P_j\langle g' \rangle\) for \(i \neq j\).
  We see from~\cref{rem:generating-morphisms} that any such map must take \(P_i\langle g \rangle\) to \(P_j \langle g+1 \rangle\).
  In other words, the differentials has the form
  \[P_i\langle g \rangle[h] \to P_j\langle g+1 \rangle[h-1]\]
  if \(i \neq j\).
  Since \(P_i\langle g \rangle[h]\) and \(P_j\langle g+1 \rangle[h-1]\) have the same level, any differential from a graded shift of \(P_i\) to a graded shift of \(P_j\) keeps the level unchanged.
  The other possibility is a map from \(P_i\langle g \rangle\) to \(P_i\langle g' \rangle\).
  Once again, it is clear from~\cref{rem:generating-morphisms} that such differentials have the form
  \[P_i \langle g \rangle[h] \to P_i\langle g' \rangle[h-1],\]
  where \(g'\) is either \(g\) or \(g + 2\).
  Any component of the form
  \[P_i \langle g \rangle[h] \to P_i\langle g \rangle[h-1]\]
  is necessarily a multiple of the identity map, and identity maps do not appear in the Khovanov--Seidel complex.
  Hence the only possibility is multiples of the loop map, which have the form
  \[P_i \langle g \rangle[h] \to P_i\langle g+2 \rangle[h-1].\]  
  Therefore any portion of the differential from a graded shift of a \(P_i\) to another graded shift of itself raises the level by one.

@@

Recall from~\cref{defprop:cohomology-filtration} that any object \(X\) of \(\mathcal{C}_m\) has a canonical cohomology filtration whose factors are homological shifts of objects in \(\mathcal{A}_m\).
For general \(X\), the cohomology filtration can be arbitrary.
However, it turns out that the cohomology filtration is highly constrained if \(X\) is spherical, and these constraints are given by wiggly pseudotriangulations, as explained in the following proposition.
\begin{proposition}\label{prop:hn-filtration}
  Let \(X \in \mathcal{C}_{n+1}\) be a spherical object.
  Consider its cohomology filtration
  \[
    \begin{tikzcd}[column sep=0.3cm]
      0 \arrow{rr} && X_{-l} \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{-k} = X \arrow{dl}\\
      & H^l(X)[-l] \arrow[dashed]{ul}{+1} && H^{l+1}(X)[-l-1] \arrow[dashed]{ul}{+1} & \cdots & H^k(X)[-k] \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]
  Consider the set of all nonzero objects \(\{H^j(X)\}\) taken without multiplicity, and let \(T\) be corresponding set of wiggly curves.
  We have the following results.
  \begin{enumerate}
  \item 
    \(T\) is a subset of a wiggly pseudotriangulation.
  \item \(T\) contains at most one of the two irrelevant arcs.
  \end{enumerate}
\end{proposition}
\begin{proof}
  The proof of this proposition is an easy consequence of analysing cohomology filtration from the point of view of arcs on the disk with marked points.
  \asilata{TODO.}
\end{proof}

>>>>

We now describe the \(t\)-structure induced by the full subcategory of \emph{linear complexes} of \(\mathcal{C}_m\); a variant of this construction is also described in~\cite[Section 2.5]{lic.que:21}.
It is convenient to think of an object \(P^{\bullet}\) of \(\mathcal{C}_m\)
as bigraded (projective) \(A_m\)-module equipped with a differential \(d \colon P^{\bullet} \to P^{\bullet}\).
More precisely, we can write
\[P^{\bullet} = \bigoplus_{g,h,i} M_i\langle g \rangle[h],\]
where \(g\) is the internal grading shift, and \(h\) is the homological grading shift, and 
each object \(M_i\) is a direct sum of copies of objects in \(\{P_1,\ldots, P_m\}\).
Set \(P^h\) to be the direct sum of all summands that have homological grading shift \(h\).
That is, in the notation above, we have
\[P^{h} = \bigoplus_{g,i} M_i\langle g \rangle[h].\]

@@

The Khovanov--Seidel correspondence (explained in Section 4 of~\cite{kho.sei:02}) associates objects of \(\mathcal{C}_{n+1}\) (up to grading shift) to curves on a configuration of \(n+2\) marked points arranged in a straight line.

@@

Let us first characterise wiggly arcs in terms of objects of \(\mathcal{C}_m\).
\begin{lemma}
  Every wiggly arc corresponds, up to grading shift, to an indecomposable spherical object in the heart \(\mathcal{A}_m\) of the standard \(t\)-structure on \(\mathcal{C}_m\).
\end{lemma}
\begin{proof}
  Let \(P^{\bullet}\) be the complex built from an arbitrary (not necessarily wiggly) arc.
  The results of~\cite{kho.sei:02} imply that \(P^{\bullet}\) is spherical and indecomposable; we now show that it belongs to \(\mathcal{A}_m\) up to grading shift if and only if the arc is a wiggly arc.

  Recall that the procedure to build up the complex of projective modules from an arc explicitly describes the differentials.    
  Each component of these differentials is a map of the form
  \[P_i\langle g \rangle[h] \to P_j\langle g' \rangle[h-1].\]
  First consider a component of the differential from some \(P_i\langle g \rangle\) to some \(P_j\langle g' \rangle\) for \(i \neq j\).
  We see from~\cref{rem:generating-morphisms} that any such map must take \(P_i\langle g \rangle\) to \(P_j \langle g+1 \rangle\).
  In other words, the differentials has the form
  \[P_i\langle g \rangle[h] \to P_j\langle g+1 \rangle[h-1]\]
  if \(i \neq j\).
  Since \(P_i\langle g \rangle[h]\) and \(P_j\langle g+1 \rangle[h-1]\) have the same level, any differential from a graded shift of \(P_i\) to a graded shift of \(P_j\) keeps the level unchanged.
  The other possibility is a map from \(P_i\langle g \rangle\) to \(P_i\langle g' \rangle\).
  Once again, it is clear from~\cref{rem:generating-morphisms} that such differentials have the form
  \[P_i \langle g \rangle[h] \to P_i\langle g' \rangle[h-1],\]
  where \(g'\) is either \(g\) or \(g + 2\).
  Any component of the form
  \[P_i \langle g \rangle[h] \to P_i\langle g \rangle[h-1]\]
  is necessarily a multiple of the identity map, and identity maps do not appear in the Khovanov--Seidel complex.
  Hence the only possibility is multiples of the loop map, which have the form
  \[P_i \langle g \rangle[h] \to P_i\langle g+2 \rangle[h-1].\]  
  Therefore any portion of the differential from a graded shift of a \(P_i\) to another graded shift of itself raises the level by one.

@@

Recall from~\cref{defprop:cohomology-filtration} that any object \(X\) of \(\mathcal{C}_m\) has a canonical cohomology filtration whose factors are homological shifts of objects in \(\mathcal{A}_m\).
For general \(X\), the cohomology filtration can be arbitrary.
However, it turns out that the cohomology filtration is highly constrained if \(X\) is spherical, and these constraints are given by wiggly pseudotriangulations.
\begin{proposition}\label{prop:hn-filtration}
  Let \(X \in \mathcal{C}_{n+1}\) be a spherical object.
  Consider its cohomology filtration
  \[
    \begin{tikzcd}[column sep=0.3cm]
      0 \arrow{rr} && X_{-l} \arrow{dl} \arrow{rr} && \cdots \arrow{rr} \arrow{dl} && X_{-k} = X \arrow{dl}\\
      & H^l(X)[-l] \arrow[dashed]{ul}{+1} && H^{l+1}(X)[-l-1] \arrow[dashed]{ul}{+1} & \cdots & H^k(X)[-k] \arrow[dashed]{ul}{+1}
    \end{tikzcd}
  \]
  Consider the set of all possible indecomposable summands of the cohomology objects \(\{H^j(X)\}\) taken without multiplicity.
  Let \(T\) be corresponding set of wiggly curves.
  We have the following results.
  \begin{enumerate}
  \item 
    \(T\) is a subset of a wiggly pseudotriangulation.
  \item \(T\) contains at most one of the two irrelevant arcs.
  \end{enumerate}
\end{proposition}
\begin{proof}
  The proof of this proposition is an easy consequence of analysing cohomology filtration from the point of view of arcs on the disk with marked points.
  \asilata{TODO.}
\end{proof}
